Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2018-07-30 paper/manuscript.tex
Mention that restraints in bound state can add a bias.

\paragraph{Structural definition of the bound state and weak binders.}
\todo[inline, color={red!40}]{AR: This paragraph presents concepts that are common to absolute and relative binding free energy calculations, but it is useful to have this \textit{before} talking about restraints. Should we move the "common" section before the "differences" section?}

In binding free energy calculations, extra care should be taken when simulating the bound state, especially when dealing with weak binders and absolute free energy calculations in combination with enhanced sampling techniques.
In principle, only configurations of the receptor-ligand complex that we consider "bound" should be sampled from the bound state, and, in atomistic simulations, this requires to establish a structural definition of the bound state.
In practice, the simulation of the bound state starts with the ligand already placed in the binding site and relies on kinetic trapping to maintain a bound complex.
However, this strategy may not be sufficient if the complex dissociation rate is high or if methodologies such as Hamiltonian replica exchange [ref] and expanded ensemble [ref] are employed in absolute free energy calculations to enhance sampling since, in both cases, the ligand may find a way out of the binding site on timescales that are achievable by modern MD simulations.
A solution commonly adopted in these cases is the use of one or more restraints making the unbound configurations energetically unfavorable through the addition of extra terms in the potential function.
It should be noted that these additional energy terms can introduce a bias in the predicted free energy that should be properly handled (see Section \ref{sec:standardstate-restraints} for details)\todo[inline, color={red!40}]{AR: Maybe point to the specific section that talks about the restraint protocol and how to remove the bias.}.
In practice, even with restraints, it is not always trivial to force a molecular dynamics simulation to explore a restricted region of the configurational space with a complicated geometry such as a binding site.
If the ligand is a tight binder\todo[inline, color={red!40}]{AR: Should we give an order of magnitude to define a tight and a weak binder?}, it is usually safe to employ a restraint that allows some of the unbound configurations to be sampled since they generally do not contribute significantly to the partition function (i.e., they have a relatively small Boltzmann weight) as long as the sampled volume is not so large that their cumulative contribution becomes significant.
However, this is more problematic when dealing with weak binders as the unbound configurations can have a non-negligible Boltzmann weight, and their binding affinity can exhibit a significant dependency on the restraint type and parameters that are used to determined the sampled volume.

@@

\todo[inline, color={red!40}]{AR: Add that with weak binders, since we use restraints to define the bound state, the predicted free energy depends on the restraint and relative calculations may need restraints as well. Also, when kinetic trapping doesn't work a restraint is required in the bound state as well, but the restraint can introduce a bias in the free energy. One possible strategy is to use a relatively soft restraint (harmonic, flat-bottom) and post-process the data to remove "unbound" configurations and the enthalpic bias introduced by the restraint.}

@@

>>>>

@@

@@

\subsubsection{Structural definition of the bound state and weak binders}

In binding free energy calculations, extra care should be taken when simulating the bound state, especially when dealing with weak binders and absolute free energy calculations in combination with enhanced sampling techniques.
In principle, only configurations of the receptor-ligand complex that we consider "bound" should be sampled from the bound state, and, in atomistic simulations, this requires to establish a structural definition of the bound state.
In practice, the simulation of the bound state starts with the ligand already placed in the binding site and relies on kinetic trapping to maintain a bound complex.
However, this strategy may not be sufficient if the complex dissociation rate is high or if methodologies such as Hamiltonian replica exchange [ref] and expanded ensemble [ref] are employed in absolute free energy calculations to enhance sampling since, in both cases, the ligand may find a way out of the binding site on timescales that are achievable by modern MD simulations.
A solution commonly adopted in these cases is the use of one or more restraints making the unbound configurations energetically unfavorable through the addition of extra terms in the potential function.
In practice, even with restraints, it is not always trivial to force a molecular dynamics simulation to explore a restricted region of the configurational space with a complicated geometry such as a binding site.
If the ligand is a tight binder\todo[inline, color={red!40}]{AR: Should we give an order of magnitude to define a tight and a weak binder?}, it is usually safe to employ a restraint that allows some of the unbound configurations to be sampled since they generally contribute negligibly to the partition function (i.e., they have a relatively small Boltzmann weight) as long as the sampled volume is not so large that their cumulative contribution becomes significant.
However, this is more problematic when dealing with weak binders as the unbound configurations can have a non-negligible Boltzmann weight, and their binding affinity can exhibit a significant dependency on the restraint type and parameters that are used to determine the sampled volume [ref].
Moreover, it should be noted that the additional potential energy terms used to model the restraints can introduce a bias in the predicted free energy.
The bias can be removed at the analysis stage through reweighting techniques, but this procedure can increase the statistical uncertainty of the binding free energy estimate when the restraints are so strong that the overlap with the reweighted state is diminished [ref].
